---
title: "Polling Simulation: Sampling and the Central Limit Theorem"
---

In this activity, you'll use data from the 2024 Cooperative Election Study (CES) — a large-scale national survey conducted during the 2024 election — to simulate what happens when a news organization tries to estimate public opinion from a poll sample. The full dataset has nearly 55,000 respondents, which we'll treat as our population.

Your goal is to see firsthand how sample size affects the accuracy of estimates, and what the distribution of sample estimates actually looks like — the core idea behind the Central Limit Theorem.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the CES data

Load the CES 2024 religion dataset from this URL and save it to a dataframe called `ces`:

`https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/ces_2024_religion.csv`

```{r}
ces <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/ces_2024_religion.csv")
```

### Task 3: Calculate the population proportion

We'll focus on one question: what share of American adults identify as religiously unaffiliated (atheist, agnostic, or nothing in particular)? This has been a major story in American religion journalism for years.

First, we'll create a new column called `unaffiliated` that is `TRUE` if a respondent's `trad` value is `"Atheist"`, `"Agnostic"`, or `"Nothing in Particular"`, and `FALSE` otherwise. Then calculate the proportion of the full dataset that is unaffiliated. Save this value as `population_proportion`.

```{r}
ces <- ces |>
  mutate(unaffiliated = trad %in% c("Atheist", "Agnostic", "Nothing in Particular"))

proportion <- mean(ces$unaffiliated)
```

That population proportion is the "true" population value that any poll is trying to estimate. Write it down — you'll be comparing everything else to it.

### Task 4: Take a single small sample

Use `set.seed(42)` and then sample 50 rows from `ces`. Save it as `sample_50`. Calculate the proportion of unaffiliated respondents in your sample.

```{r}
set.seed(42)
sample_50 <-  ces |> sample_n(50)
mean(sample_50$unaffiliated)
```

How close is your sample estimate to the original proportion? Is it surprising?

This sample estimate is decently close to the original proportion. It could be more accurate, but since this is a small sample size I am not surprised by its closeness to the original proportion.

### Task 5: Take a single larger sample

Now sample 500 rows. Set a different seed value, grab the sample and save it as `sample_500`. Calculate the proportion of unaffiliated respondents in this sample.

```{r}
set.seed(405)
sample_500 <-  ces |> sample_n(500)
mean(sample_500$unaffiliated)

```

How did the estimate change compared to the sample of 50? What does this suggest about sample size?

This estimate got larger compared to the sample of 50, making it more accurate and closer to the original estimate. This shows that a larger sample size tends to get closer to representing the population. 

### Task 6: Simulate many polls

A single sample can get lucky or unlucky. To understand how polling actually works, we need to see the *distribution* of estimates across many samples. The code below draws 1,000 samples of size 50, calculates the proportion of unaffiliated respondents in each, and stores the results.

```{r}
set.seed(42)

simulate_polls <- function(data, sample_size, n_simulations = 1000) {
  proportions <- replicate(n_simulations, {
    s <- data |> sample_n(sample_size)
    mean(s$unaffiliated)
  })
  tibble(proportion = proportions, sample_size = sample_size)
}

simulations_50 <- simulate_polls(ces, 50)
simulations_500 <- simulate_polls(ces, 500)
```

### Task 7: Visualize the simulation results

Create two histograms side by side — one for the 1,000 samples of size 50 and one for the 1,000 samples of size 500. Add a vertical line at `proportion` to each so you can see how well the samples are centering on the true value.

```{r}
simulations_all <- bind_rows(simulations_50, simulations_500) |> 
  mutate(sample_size = paste("n =", sample_size))

ggplot(simulations_all, aes(x = proportion)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  geom_vline(xintercept = proportion, color = "red", linetype = "dashed", linewidth = 1) +
  facet_wrap(~sample_size) +
  labs(
    title = "Distribution of Sample Estimates: Share Religiously Unaffiliated",
    subtitle = "Red line = true population proportion. Each histogram shows 1,000 simulated polls.",
    x = "Sample Proportion",
    y = "Count",
    caption = "Source: 2024 Cooperative Election Study"
  ) +
  theme_minimal()
```

### Task 8: Reflect on what you see

**What shape do both histograms have?** Describe them in plain language. Does it surprise you that this shape emerges from a variable that is simply TRUE or FALSE for each respondent?

Both of these histograms have a symmetrical shape and are normally distributed. I am not surprised this shape emerged from a true or false questionairre because there aren't as many options for outliers or random data points. 

**Compare the two histograms.** What is the most visible difference between the n=50 and n=500 distributions? What does that difference tell you about what happens when a poll uses a larger sample?

The biggest difference between these two histograms is that the one that uses the n=50 sample size is less concentrated around a central point. This shows that when a poll uses a larger sample, you will be able to observe more uniform data and start to approach the main statistical figure that represents the population. 

**Both distributions are centered near the red line.** What does it mean that 1,000 different random samples — each one drawn independently — tend to cluster around the same value? What would it mean for journalism if that *weren't* the case?

Clustering around the same value shows that there's some kind of consensus among the population. If this were not the case, that would show that there is a lot of variation in answers and might not be enough information to base a main argument on a group consensus off of. 

**A reporter tells you their story is based on a poll of 50 people.** Based on what you've seen in the simulation, what questions would you want to ask before trusting that number? You don't need to use any technical terms — just describe what you observed.

I would ask how many people are part of the population they are surveying. I would want them to perhaps take a larger sample size if possible, but if not I would just caution them of using those statistics as end all be all arguments for their article. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
