---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
install.packages("janitor")
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |> mutate(violent_rate_2019 = violent2019/pop2019 * 1000, property_rate_2019 = property2019/pop2019 * 1000, violent_rate_2020 = violent2020/pop2020 * 1000, property_rate_2020 = property2020/pop2020 * 1000) 
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

Ocean City has the highest rate of property and violent crime alongside a significant jump up in crime from 2019 to 2020. Interestingly, Baltimore City, which is much larger than Ocean City, had a higher rate in violent crime in 2019 before dropping in 2020. Property crime in Baltimore is lower than in Ocean City both years, with Ocean City nearly double and triple the Baltimore property crime rate each year. I think this is probably because of how many tourists visit Ocean City who likely fall victim to robberies or theft at a higher rate than locals. I'm also interested to know how Ocean City's rate of violent crime stayed so high in 2020 during COVID, too. 


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate(violent_percent_change = (`violent2020`- `violent2019`)/`violent2019` * 100, property_percent_change = (`property2020`-`property2019`)/`property2019`* 100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

Based on this data, it seems that Ocean City is an outlier. COVID seemed to bring property crime and violent crime down dramatically in most Maryland cities. I feel like that makes Ocean City the more interesting story here. How did crime increase there during COVID, and decrease nearly everywhere else? The answer to why crime decreased everywhere else is that people were inside because of lockdown mandates. Did Ocean City have different rules regarding COVID procedures? 

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Nearly every city in Maryland saw a decrease in property and violent crime during the COVID-19 pandemic, except for Ocean City. 

Ocean City, Maryland, racked up higher rates of violent crime and property crime between 2019 and 2020, outpacing Baltimore and Frederick, Maryland's most populous cities. Baltimore was closer to Ocean City's rates than Frederick, yet still, Ocean City nearly tripled the property crime rate in Baltimore in 2020, and almost doubled its violent crime rate. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.

### Task 10

Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile?

```{r}
md_cities_percentiles <- md_cities_rates |> mutate(percentile = ntile(md_cities_rates$violent_rate_2020, n=100))
```
There are 11 cities at or above the 90th percentile. 

Calculate the Interquartile Range of the 2020 violent crime rate using `summarize()` and R's `IQR()` function. The syntax is similar to how `mean()` works.

```{r}
md_cities_percentiles |> summarize(IQR = IQR(violent_rate_2020)) 
```


Do the results change your best story idea from the previous assignment? Why or why not?

I think that these results only emphasize the original story idea I had for investigating Ocean City's high crime rates. This is a high IQR, showing that there's a big gap between cities at the top of the scale versus lower down. The only thing about this new percentile table is that I might tweak my nut graf where I claim a giant difference between Ocean City and Baltimore. They are only one percentile point away from each other, so I would keep that Ocean City is still the highest in terms of these two measures of crime rates, but might lose the part about how significantly ahead of Baltimore Ocean City is. 
